---
title: "Assignment3"
author: "RF"
date: "3/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Multivariate linear models

In this assignment you will investigate the relation between different aspects of IQ and symptom severity in ASD (as measured by ADOS). The assignment is meant to make you practice on linear models, multiple predictors and interactions.

The data you will use is the outcome of the psychological testing of the children you analyzed in methods 3.
Data: https://www.dropbox.com/s/hom2qnmp9hyfwmu/Assignment3Data.csv?dl=0
The data consists of ChildID, gender, age, diagnosis (ASD), symptom severity (ADOS), several aspects of IQ (NonVerbal/Perceptual,Verbal,Social) and the visit at which the tests were administered. The questions should be answered by focusing on visit 1.

The questions you will have to answer are:

1. Assess the relation between symptom severity and IQ (focus on visit 1 and children with ASD) and report the model, a plot of the model, a couple of lines describing the quality of the model and interpreting the results. P.S. Should you scale? P.P.S. Don't forget to motivate your priors. P.P.P.S. At least one plot for results and a plot for quality of each model (here and in the next questions) would be appreciated.
1.1. Verbal IQ and ADOS
1.2. Non Verbal IQ and ADOS
1.3. Social IQ and ADOS

```{r}
library(pacman)
p_load(rethinking, dplyr, ggplot2, stringr)

df = read.csv("Assignment3Data.csv")
ASD_df = subset(df, Visit == 1 & ASD == 1)

rescalelist = c("NonVerbalIQ","VerbalIQ","SocialIQ","ADOS") #making a list with variales we want to scale
ASD_df.s = ASD_df[, colnames(ASD_df) %in% rescalelist] %>% #telling R to select the colnames which is both in ASD_df and rescalelist
  lapply(.,function(x) scale(x,center= mean(x,na.rm = T), scale = sd(x, na.rm = T)))%>% #applies the scale function to the columns, which is specified above.
  cbind(.,ASD_df[,! colnames(ASD_df) %in% rescalelist]) #bind the scaled columns with the remaining columns in ASD_df to make dataframe ASD_df.s

#Motivate priors - expected mean and sigma for alpha and beta - based on our knowlegde
  #Need an explanation for this

#Fitting some models
m1 <- map(
  alist(
    ADOS ~ dnorm(mu,sigma),
    mu <- a + b*VerbalIQ,
    a ~ dnorm(0,0.5), 
    b ~ dnorm(0,0.5),
    sigma ~ dunif(0,2)
  ),data = ASD_df.s)

m2 <- map(
  alist(
    ADOS ~ dnorm(mu,sigma),
    mu <- a + b*NonVerbalIQ,
    a ~ dnorm(0,0.5), 
    b ~ dnorm(0,0.5),
    sigma ~ dunif(0,2)
  ),data = ASD_df.s)


m3 <- map(
  alist(
    ADOS ~ dnorm(mu,sigma),
    mu <- a + b*SocialIQ,
    a ~ dnorm(0,0.5), 
    b ~ dnorm(0,0.5),
    sigma ~ dunif(0,2)
  ),data = ASD_df.s)

precis(m3)
precis(m2)
precis(m1)

#Motivate the priors, alpha, beta and sigma
#Naive prior due to correlation in both direction (ados is both negative and postitive). If there is an increase in IQ of std we expect an increase in ADOS of 0 which might vary with a std 0.5. Sigma has to be bounded at zero. 

#Make some pretty plots

#Kenneth made a pretty, pretty function <3 

gggplot = function(model, x = NULL , y = NULL, dataframe){ 
  #where:
    #model is a model resulting from MAP
    #x is the x variable as a string (can be left unspecified)
    #y is the y variable as a string (can be left unspecified)
    #dataframe is the dataframe from which the model is trained upon
  
    #requires packages: dplyr, ggplot2, stringr and rethinking
  
  if (is.null(y) == T){
      #if x isn't specified - extract y
    temp <- flist_untag(model@formula)[[1]]
    y <- as.character(temp[[2]]) 
  }
  if (is.null(x) == T){
      #if x isn't specified - extract x
    temp <- flist_untag(model@formula)[[2]]
    x <- gsub(" ","",gsub(".*\\*","",as.character(temp[[3]])[3]))
  }
    
    #extracting data from the dataframe
  xd <- dplyr::select(dataframe, x)
  yd <- dplyr::select(dataframe, y)
  
    #calculate the prior data
  post <- extract.samples(model) #posterior from samples 
  mu.link <- function(x) post$a + post$b*x
  VIQ.seq <- seq( from=min(xd) , to=max(xd) , by=0.1 )
  mu <- sapply( VIQ.seq , mu.link )
  mu.mean <- apply( mu , 2 , mean )
  mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )
  temp_list = list(VIQ.seq); names(temp_list) <- x #naming the vector in the list (since the sim function requires it)
  sim.VIQ <- sim( model , data=temp_list )
  VIQ.PI <- apply( sim.VIQ , 2 , PI , prob=0.89 )
  
    #transform the data from ggplot
      #for making the line
  temp_df <- data.frame(VIQ.seq = VIQ.seq, mu.mean = mu.mean) 
      #for making the HDPI
  temp_matrix <- t(data.frame(mu.HPDI = mu.HPDI)) #t() is a tranpose function which returns a matrix
  temp_df1 = data.frame(lower = temp_matrix[,1], upper = temp_matrix[,2], 
                        mu.mean = mu.mean, VIQ.seq = VIQ.seq)
      #for making the PI
  temp_matrix <- t(data.frame(VIQ.PI = VIQ.PI)) 
  temp_df2 = data.frame(lower = temp_matrix[,1], upper = temp_matrix[,2], 
                        mu.mean = mu.mean, VIQ.seq = VIQ.seq)
      #for the points
  dataframe = data.frame(xd, yd)
    #taking the data to new column as to chose the name (maintaining old name for the names on the plot)
  dataframe$c1 = dataframe[,1]
  dataframe$c2 = dataframe[,2]
  
    #make the plot
  grob = ggplot(dataframe, aes(x = c1, y = c2)) +
    geom_ribbon(data = temp_df2,
                aes(x = VIQ.seq, y = mu.mean, ymin = lower, ymax = upper),
                fill = "grey70", alpha = 2/3) +
      geom_ribbon(data = temp_df1,
                aes(x = VIQ.seq, y = mu.mean, ymin = lower, ymax = upper),
                fill = "grey83", alpha = 2/3) +
    geom_line(data = temp_df, aes(x = VIQ.seq, y = mu.mean))+
    geom_point(color = "navyblue", shape = 1, size = 1.5, alpha = 2/3)+
    xlab(colnames(dataframe)[1]) + ylab(colnames(dataframe)[2]) #using the names for the axis
  return(grob)
}  #go go gadget fancy plot

plot.m1 = gggplot(m1,dataframe = ASD_df.s)
plot.m1

plot.m2 = gggplot(m2,dataframe = ASD_df.s)
plot.m2

plot.m3 = gggplot(m3, dataframe = ASD_df.s)
plot.m3


#Plots for assessing model quality? Can we use the one above?
```

2. Do the different aspects of IQ account for different portions of the variance in ADOS? 
2.1. Does it make sense to have all IQ measures in the same model? First write a few lines answering the question and motivating your answer, including a discussion as to what happens when you put all of them in the same model. Then build a model following your answer. If your answer is "no", you are not free, you still have to answer: are there alternative ways of answering the question?
2.2. Build the model, assess its quality, write a few lines interpreting the results.

```{r}
#2.1
#Including more predictors = more variance explained 
#BUT! one should be vary of correlation/multicollinariety, i.e. the predictors contain the same information. 
#Check for correlation between predictors/use pairs plot from the book(?)
cor(ASD_df.s)

#Including or not including?
#Alternative ways?
#Model building is rouhgly the same procedure as Q1

```

3. Let's now include also the TD children. Does it make sense to ask whether IQ and ADOS are related? Motivate your answer. In any case, if you wanted to build a model to answer that question, which model would you build? Run the model, assess its quality, write a few lines interpreting the results.

4. Let's discuss contents:
4.1. You have three scores for IQ, do they show shared variance? Is that the same in TD and ASD? What does that tell us about IQ?
4.2. You have explored the relation between IQ and ADOS. How do you explain that relation from a cognitive perspective? N.B. You can present alternative hypotheses.

5. Bonus questions: Including measurement errors. 
5.1. Let's assume that ADOS has a measurement error of 1. How would you model that in any of the previous models? 
5.2. We know that IQ has an estimated measurement error of 2.12. How would you include that? 


