---
title: "Computational Modeling - Week 4 - Assignment 2 - Part 1"
author: "Riccardo Fusaroli"
date: "2/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results
- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)

```{r}
library(pacman)
p_load(rethinking,ggplot2)


length = 1e4
p_grid = seq(0,1, length.out = length)
Uni.prior = rep(1,length)

#Plotting prior
temp_df = data.frame(grid = p_grid, prior = Uni.prior)
prior = ggplot(temp_df, aes(grid,prior))+geom_line()+xlab("P_grid")+ylab("Uniform Prior")
prior

#Calculating likelihood for Riccardo based on flat prior + posterior probability. 
RF.like = dbinom(3,size = 6, prob = p_grid)
unstd.posterior = RF.like*Uni.prior
RF.PostUni = unstd.posterior/sum(unstd.posterior)

#Is Riccardo better than chance?
sum(RF.PostUni[p_grid > 0.5]) 
temp_df = data.frame(grid = p_grid, posterior = RF.PostUni)

ggplot(temp_df, aes(grid,posterior))+geom_line()+ xlab("Percentage correct answers")+ylab("Posterior probability")


#1.2 - quadratic approx.
#Explaining the code: map() needs a list with defined values/parameter etc. Here, we need to define the prior p and the likelihood of c (correct answers). p is a uniform prior (why we use dunif) and c is binomial (why we use dbinom).
know.qa = map(
  alist(
    c ~ dbinom(6,p),
    p ~ dunif(0,1)
  ),
  data=list(c=3))

precis(know.qa)
```

2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
2a. Produce plots of the prior, and posterior for each teacher.

```{r}
#Making a df with results
results_df = data.frame(teacher = c("RF","KT","JS","MW"), correct = c(3,2,160,66), n_quest = c(6,2,198,132))

#Looping through the teachers
for (teacher in results_df$teacher){
  correct = results_df$correct[results_df$teacher == teacher]
  size = results_df$n_quest[results_df$teacher == teacher]
  likelihood = dbinom(correct,size = size,prob = p_grid)
  unstd.posterior = likelihood*Uni.prior
  posterior = unstd.posterior/sum(unstd.posterior)
  temp_df = data.frame(grid = p_grid, posterior = posterior)
  temp_plot = ggplot(temp_df,aes(grid,posterior))+geom_line()+xlab("probability of correct answer")+ylab("posterior probability")
  assign(paste(teacher, "p",sep =""),temp_plot)
  assign(paste(teacher, ".PostUni", sep = ""), posterior)
}

temp_df = data.frame(p_grid = rep(p_grid, 4), 
                      posterior = c(MW.PostUni, RF.PostUni, KT.PostUni, JS.PostUni), 
                      teacher = c(rep("MW", length(MW.PostUni)), rep("RF", length(RF.PostUni)),
                                  rep("KT", length(KT.PostUni)), rep("JS", length(JS.PostUni)))
                      )

ggplot(temp_df, aes(p_grid, posterior, color = teacher))+ geom_line()+xlab("Percentage correct answers")+ylab("Posterior probability")
ggplot(subset(temp_df, teacher == "MW"), aes(p_grid, posterior))+geom_line()+ xlab("Percentage correct answers")+ylab("Posterior probability")
ggplot(subset(temp_df, teacher == "KT"), aes(p_grid, posterior))+geom_line()+ xlab("Percentage correct answers")+ylab("Posterior probability")
ggplot(subset(temp_df, teacher == "JS"), aes(p_grid, posterior))+geom_line()+ xlab("Percentage correct answers")+ylab("Posterior probability")
ggplot(subset(temp_df, teacher == "RF"), aes(p_grid, posterior))+geom_line()+ xlab("Percentage correct answers")+ylab("Posterior probability")

#Balancing uncertainty and most probable values
#Criteria - who to chose - are we looking for a superstar? (Kristian) Or are we more conservative? 

```


3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.

```{r}
#Making da new prior - we expect more of these smart guys!
In.prior = dnorm(p_grid, mean = 0.8, sd = 0.2)

#Plotting prior
temp_df = data.frame(grid = p_grid, prior = In.prior)
ggplot(temp_df,aes(grid,prior))+geom_line()

#What does this prior tells us?
  #The tail is skewed --> higher probability assigned to higher scores
  #Remember: the posterior is calculated as the product between the likelihood and the prior (so ofc, if prior changes, the posterior changes shape)

  #The posterior distribution for the 4 professors moves towards a higher probability for higher values of p (the parameter - in this case, correct answers --> a measure of cog sci knowlegde) 

for (teacher in results_df$teacher){
  correct = results_df$correct[results_df$teacher == teacher]
  size = results_df$n_quest[results_df$teacher == teacher]
  likelihood = dbinom(correct,size = size,prob = p_grid)
  unstd.posterior = likelihood*In.prior
  posterior = unstd.posterior/sum(unstd.posterior)
  temp_df = data.frame(grid = p_grid, posterior = posterior)
  temp_plot = ggplot(temp_df,aes(grid,posterior))+geom_line()+xlab("probability of correct answer")+ylab("posterior probability")
  assign(paste(teacher, "p",sep =""),temp_plot)
  assign(paste(teacher, ".PostIn", sep = ""), posterior)
}

temp_df = data.frame(p_grid = rep(p_grid, 4), 
                      posterior = c(MW.PostIn, RF.PostIn, KT.PostIn, JS.PostIn),
                      posterior.Uni = c(MW.PostUni, RF.PostUni, KT.PostUni, JS.PostUni),
                      teacher = c(rep("MW", length(MW.PostIn)), rep("RF", length(RF.PostIn)),
                                  rep("KT", length(KT.PostIn)), rep("JS", length(JS.PostIn)))
                      )

ggplot(temp_df, aes(p_grid, posterior, color = teacher))+ geom_line()+xlab("Percentage correct answers")+ylab("Posterior probability")

ggplot()+geom_line(data = subset(temp_df, teacher == "MW"), aes(p_grid,posterior.Uni), color = "black")+geom_line(data = subset(temp_df, teacher == "MW"), aes(p_grid,posterior), color = "red")+ xlab("Percentage correct answers")+ylab("Posterior probability")


ggplot()+geom_line(data = subset(temp_df, teacher == "KT"), aes(p_grid,posterior.Uni), color = "black")+geom_line(data = subset(temp_df, teacher == "KT"), aes(p_grid,posterior), color = "red")+ xlab("Percentage correct answers")+ylab("Posterior probability")

ggplot()+geom_line(data = subset(temp_df, teacher == "RF"), aes(p_grid,posterior.Uni), color = "black")+geom_line(data = subset(temp_df, teacher == "RF"), aes(p_grid,posterior), color = "red")+ xlab("Percentage correct answers")+ylab("Posterior probability")

ggplot()+geom_line(data = subset(temp_df, teacher == "JS"), aes(p_grid,posterior.Uni), color = "black")+geom_line(data = subset(temp_df, teacher == "JS"), aes(p_grid,posterior), color = "red")+ xlab("Percentage correct answers")+ylab("Posterior probability")

```

4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?

```{r}
#Multiplying data with 100
results_df$correctx100 = results_df$correct*100
results_df$n_questx100 = results_df$n_quest*100

#With uniform/flat prior
for (teacher in results_df$teacher){
  correct = results_df$correctx100[results_df$teacher == teacher]
  size = results_df$n_questx100[results_df$teacher == teacher]
  likelihood = dbinom(correct,size = size,prob = p_grid)
  unstd.posterior = likelihood*Uni.prior
  posterior = unstd.posterior/sum(unstd.posterior)
  temp_df = data.frame(grid = p_grid, posterior = posterior)
  temp_plot = ggplot(temp_df,aes(grid,posterior))+geom_line()+xlab("probability of correct answer")+ylab("posterior probability")
  assign(paste(teacher, "p",sep =""),temp_plot)
  assign(paste(teacher, ".PostUnix100", sep = ""), posterior)
}

temp_df = data.frame(p_grid = rep(p_grid, 4), 
                      posterior = c(MW.Postx100, RF.Postx100, KT.Postx100, JS.Postx100), 
                      teacher = c(rep("MW", length(MW.Postx100)), rep("RF", length(RF.Postx100)),
                                  rep("KT", length(KT.Postx100)), rep("JS", length(JS.Postx100)))
                      )

ggplot(temp_df, aes(p_grid, posterior, color = teacher))+ geom_line()+xlab("Percentage correct answers")+ylab("Posterior probability")

#With informed prior
for (teacher in results_df$teacher){
  correct = results_df$correctx100[results_df$teacher == teacher]
  size = results_df$n_questx100[results_df$teacher == teacher]
  likelihood = dbinom(correct,size = size,prob = p_grid)
  unstd.posterior = likelihood*In.prior
  posterior = unstd.posterior/sum(unstd.posterior)
  temp_df = data.frame(grid = p_grid, posterior = posterior)
  temp_plot = ggplot(temp_df,aes(grid,posterior))+geom_line()+xlab("probability of correct answer")+ylab("posterior probability")
  assign(paste(teacher, "p",sep =""),temp_plot)
  assign(paste(teacher, ".PostInx100", sep = ""), posterior)
}

temp_df = data.frame(p_grid = rep(p_grid, 4), 
                      posterior = c(MW.PostInx100, RF.PostInx100, KT.PostInx100, JS.PostInx100), 
                      teacher = c(rep("MW", length(MW.PostInx100)), rep("RF", length(RF.PostInx100)),
                                  rep("KT", length(KT.PostInx100)), rep("JS", length(JS.PostInx100)))
                      )

ggplot(temp_df, aes(p_grid, posterior, color = teacher))+ geom_line()+xlab("Percentage correct answers")+ylab("Posterior probability")

#The proability distribution narrows with more data = the number of possible values of p (the parameter) becomes smaller - the contrast between probable numbers and unprobable numbers becomes bigger. 
```

5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?

```{r}
#Making a conservative prior -> we expect them to score 0 (way below chance) --> assigning zero probability to values above chance (0.5) --> skewing prior probability = skewing posterior probability
#Following a gaussian distribution, centered around 0.5, but what about SD?
```

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.
```{r}
library(brms)
d <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

FlatModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("uniform(0,1)", class = "Intercept"))
plot(FlatModel)
PositiveModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.8,0.2)", class = "Intercept"))
plot(PositiveModel)
SkepticalModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.5,0.01)", class = "Intercept"))
plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?

###Second part: Focusing on predictions

Questions to be answered (but see guidance below):
1- Write a paragraph discussing how assessment of prediction performance is different in Bayesian vs. frequentist models
2- Provide at least one plot and one written line discussing prediction errors for each of the teachers.

This is the old data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

This is the new data:
- Riccardo: 9 correct answers out of 10 questions (then he freaks out about teaching preparation and leaves)
- Kristian: 8 correct answers out of 12 questions
- Josh: 148 correct answers out of 172 questions (again, Josh never gets bored)
- Mikkel: 34 correct answers out of 65 questions

Guidance Tips

1. There are at least two ways of assessing predictions.
2. Last year's results are this year's expectations.
3. Are the parameter estimates changing? (way 1)
4. How does the new data look in last year's predictive posterior? (way 2)

```{r}
#Adding new data to result_df
results_df$correct.new = c(9,8,148,34)
results_df$n_quest.new = c(10,12,172,65)

#We assume the an informed prior
new.Prior = data.frame(RF = RF.PostIn, MW = MW.PostIn, KT = KT.PostIn, JS = JS.PostIn)

#What do we want the loop to do? 1) go through all teachers, and 2) calculate likelihood, 3) calculate unstd.posterior and 3) calculate poserior. We want to plot and save posterior. 
for (teacher in results_df$teacher){
  correct = results_df$correct.new[results_df$teacher == teacher]
  size = results_df$n_quest.new[results_df$teacher == teacher]
  prior = select(new.Prior, teacher)[,1] 
  likelihood = dbinom(correct, size = size, prob = p_grid)
  unstd.posterior = likelihood*prior
  posterior = unstd.posterior/sum(unstd.posterior)
  temp_df = data.frame(p_grid = rep(p_grid, 2), 
                        value = c(posterior, prior), 
                        type = c(rep("posterior", length(posterior)), rep("prior", length(prior)))
                        )
  tempgrob = ggplot(temp_df, aes(x = p_grid, y = value, color = type)) + geom_line() + ylim(0,0.0022) + ggtitle(teacher)
  assign(paste("n", teacher, "p", sep = ""), tempgrob)
}

likelihood_mikkel = dbinom(66,size = 132, prob = p_grid)
likelihood_josh = dbinom(160,size = 198, prob = p_grid)
likelihood_kristian = dbinom(2,size = 2, prob = p_grid)


unstd.posterior = likelihood_mikkel*uniform_prior
posterior_mikkel = unstd.posterior/sum(unstd.posterior)
prior_mikkel = posterior_mikkel

unstd.posterior = likelihood_josh*uniform_prior
posterior_josh = unstd.posterior/sum(unstd.posterior)
prior_josh = posterior_josh

unstd.posterior = likelihood_kristian*uniform_prior
posterior_kristian = unstd.posterior/sum(unstd.posterior)
prior_kristian = posterior_kristian

plot(prior_josh)
plot(prior_mikkel)
plot(prior_kristian)

likelihood_mikkel_new = dbinom(34,size = 65, prob = p_grid)
unstd.posterior = likelihood_mikkel_new*prior_mikkel
posterior_new_mikkel = unstd.posterior/sum(unstd.posterior)

likelihood_josh_new = dbinom(148,size = 172 , prob = p_grid)
unstd.posterior = likelihood_josh_new*prior_josh
posterior_new_josh = unstd.posterior/sum(unstd.posterior)

likelihood_kristian_new = dbinom(8,size = 12, prob = p_grid)
unstd.posterior = likelihood_kristian_new*prior_kristian
posterior_new_kristian = unstd.posterior/sum(unstd.posterior)

plot(posterior_new_mikkel)
plot(posterior_new_josh)
plot(posterior_new_kristian)

df_oldData = data.frame(grid = rep(p_grid,4), posterior = c(posterior_riccardo,  posterior_josh, posterior_mikkel, posterior_kristian), teacher = c(rep("RF", length(p_grid)),rep("JS",length(p_grid)),rep("MW",length(p_grid)),rep("KT",length(p_grid))))

df_newData = data.frame(grid = rep(p_grid,4), posterior = c(posterior_new_riccardo, posterior_new_josh, posterior_new_mikkel, posterior_new_kristian), teacher = c(rep("RF", length(p_grid)),rep("JS",length(p_grid)),rep("MW",length(p_grid)),rep("KT",length(p_grid))))

ggplot(df_newData,aes(grid,posterior,color=teacher))+geom_line()+xlab("probability of correct answer")+ylab("posterior probability")

#Lets sample from the posterior!
samples_riccardo = sample(p_grid, prob = posterior_riccardo, size = 1e4, replace = TRUE)
samples_mikkel = sample(p_grid, prob = posterior_mikkel, size = 1e4, replace = TRUE)
samples_josh = sample(p_grid, prob = posterior_josh, size = 1e4, replace = TRUE)
samples_kristian = sample(p_grid, prob = posterior_kristian, size = 1e4, replace = TRUE)

#Making the predictive posterior - how does it hold against the actually observed data?
simplehist(rbinom(1e4, size = 10, prob = samples_riccardo))
simplehist(rbinom(1e4, size = 65, prob = samples_mikkel))
simplehist(rbinom(1e4, size = 172, prob = samples_josh))
simplehist(rbinom(1e4, size = 12, prob = samples_kristian))

#4. How does the new data look in last year's predictive posterior? (way 2)
#creating a df containing the samples 

samples_df <- data.frame(RF = sample(p_grid, size = length(p_grid), replace = T, prob = posterior_riccardo), 
                         MW = sample(p_grid, size = length(p_grid), replace = T, prob =posterior_mikkel), 
                         JS = sample(p_grid, size = length(p_grid), replace = T, prob = posterior_josh), 
                         KT = sample(p_grid, size = length(p_grid), replace = T, prob = posterior_kristian))




#make the predictive posterior from the samples

for (teacher in df_oldData$teacher){
  n_quest = df_oldData$n_quest.new[df_oldData$teacher == teacher] #extracting number of questions
  correct = df_oldData$correct.new[df_oldData$teacher == teacher]
  samples <- select(samples_df, teacher)[,1]
  if (teacher == df_oldData$teacher[1]){ # if it is the first teacher
    Predpos_df <- data.frame(
      valueAbs = rbinom(length(p_grid), n_quest, prob = samples),
      acc = rbinom(length(p_grid), n_quest, prob = samples)/n_quest,
      valueCentered = rbinom(length(p_grid), n_quest, prob = samples)-correct,
      valueRelative = (rbinom(length(p_grid), n_quest, prob = samples)-correct)/n_quest, teacher = teacher)
  } else {
    Predpos_df <- rbind(Predpos_df, 
                         Predpos_df <- data.frame(
                           valueAbs = rbinom(length(p_grid), n_quest, prob = samples),
                           acc = rbinom(length(p_grid), n_quest, prob = samples)/n_quest,
                           valueCentered = rbinom(length(p_grid), n_quest, prob = samples)-correct,
                           valueRelative = (rbinom(length(p_grid), n_quest, prob =
                                                     samples)-correct)/n_quest, teacher = teacher)
                         )
  }
}

ggplot(Predpos_df, aes(x = valueAbs, color = teacher)) + geom_density()
ggplot(Predpos_df, aes(x = acc, color = teacher)) + geom_density()
ggplot(Predpos_df, aes(x = valueCentered, color = teacher)) + geom_density()
ggplot(Predpos_df, aes(x = valueRelative, color = teacher)) + geom_density()
```

Questions to be answered (but see guidance below):
1- Write a paragraph discussing how assessment of prediction performance is different in Bayesian vs. frequentist models
2- Provide at least one plot and one written line discussing prediction errors for each of the teachers.

This is the old data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

This is the new data:
- Riccardo: 9 correct answers out of 10 questions (then he freaks out about teaching preparation and leaves)
- Kristian: 8 correct answers out of 12 questions
- Josh: 148 correct answers out of 172 questions (again, Josh never gets bored)
- Mikkel: 34 correct answers out of 65 questions

Guidance Tips

1. There are at least two ways of assessing predictions.
2. Last year's results are this year's expectations.
3. Are the parameter estimates changing? (way 1)
4. How does the new data look in last year's predictive posterior? (way 2)


### Depending on time: Questions from the handbook
2H1. Suppose there are two species of panda bear. Both are equally common in the wild and live in the same places. They look exactly alike and eat the same food, and there is yet no genetic assay capable of telling them apart. They differ however in their family sizes. Species A gives birth to twins 10% of the time, otherwise birthing a single infant. Species B births twins 20% of the time, otherwise birthing singleton infants. Assume these numbers are known with certainty, from many years of field research.
Now suppose you are managing a captive panda breeding program. You have a new female panda of unknown species, and she has just given birth to twins. What is the probability that her next birth will also be twins?

2H2. Recall all the facts from the problem above. Now compute the probability that the panda we have is from species A, assuming we have observed only the first birth and that it was twins.

2H3. Continuing on from the previous problem, suppose the same panda mother has a second birth and that it is not twins, but a singleton infant. Compute the posterior probability that this panda is species A.

2H4. A common boast of Bayesian statisticians is that Bayesian inference makes it easy to use all of the data, even if the data are of different types. So suppose now that a veterinarian comes along who has a new genetic test that she claims can identify the species of our mother panda. But the test, like all tests, is imperfect. This is the information you have about the test:
- The probability it correctly identifies a species A panda is 0.8.
- The probability it correctly identifies a species B panda is 0.65.
The vet administers the test to your panda and tells you that the test is positive for species A. First ignore your previous information from the births and compute the posterior probability that your panda is species A. Then redo your calculation, now using the birth data as well.



